\documentclass{article}
\usepackage{assignment}

\title{MAT4170 -- Assignment 3}
\author{Jon-Magnus Rosenblad}
\date{}

\lstset{
    language=Octave,
    frame=single,
    breaklines=true,
    title=\lstname,
}


\begin{document}
\maketitle

\section*{Problem 1 (3.2)}
Let $\mathbf t = (0,0,0,1,3,4,5,5)$.
The existence of the four cubic splines $\{B_{i,3}\}_{i=1}^4$ is due to our regularization not exceeding the degree of our splines.
There are $8 - (3 + 1) = 4$ of them because we have $8$ knots.
Their linear independence on $[1,3]$ follow from lemma 3.8.

The splines $\hat B_{1,3}$ and $\hat B_{2,3}$ are naturally identified with $B_{2,3}$ and $B_{3,3}$ respectively,
so they must also be linearly independent as a subset of a linearly independent set.

\section*{Problem 2 (4.5)}
Lemma 4.2 states that if $\mathbf \tau$ is a knot vector with at least $d + 2$ knots,
and it's knots occur as a subsequence of another knot vector $\mathbf t$,
then $\mathbb S_{d,\mathbf\tau}\subset \mathbb S_{d,\mathbf t}$.

We have already proven the case when $\mathbf \tau$ is $d + 1$-regular, and by extension $\mathbf t$ must also be at least $d + 1$-regular,
and we will use this in our proof of the general case.

Let $\mathbf \tau^\prime$ and $\mathbf t^\prime$ denote the $d + 1$-regularizations of $\mathbf \tau$ and $\mathbf t$ respectively.
We have a natural embedding of $\mathbb S_{d, \mathbf \mu}\subset \mathbb S_{d, \mathbf \mu^\prime}$ for $\mathbf \mu=\mathbf \tau, \mathbf t$
given by padding with zeros at the ends of the coefficient vectors.
By our restricted variant of the lemma we may take any $f\in\mathbb S_{d, \mathbf \tau}$,
lift it to $\mathbb S_{d, \mathbf \tau^\prime}$ and extend it to $\mathbb S_{d, \mathbf t^\prime}$.
It remains to show that we may restrict it again to $\mathbb S_{d, \mathbf t}$.

By the construction of our extensions $\mathbb S_{d,\mathbf\mu^\prime}\supset \mathbb S_{d, \mathbf \mu}$,
$f$ would not obtain any nonzero coefficients on our regularization,
and as no knot is inserted between the knots added in the regularization in the embedding $\mathbb S_{d,\mathbf \tau^\prime}\subset \mathbb S_{d,\mathbf t^\prime}$,
we would not obtain in any nonzero coefficients outside the span of our original coefficients.
It follows that the projection of $f$ from $\mathbb S_{d,\mathbf t^\prime}\to \mathbb S_{d,\mathbf t}$ does not alter $f$.

\section*{Problem 3 (4.6)}

As in the previous exercise, the trick is to use the regularized version of the theorem to prove the general version
through truncating the coefficient matrices.
Let $f\in \mathbb S_{d,\mathbf \tau}$ be defined by $f=\sum_j c_j B_{j,d,\tau}$.
We may express it in the $d + 1$-regularized spline-basis as $f=\sum_{j=d + 1}^{d + \# \mathbf \tau} c_{j - d} \hat B_{j,d,\tau^\prime}$.
For prettier indexing, we let $\hat c_j=c_{j - d}$ for $j=d + 1,\ldots, d + \# \mathbf \tau$ and $0$ otherwise.
Then $f=\sum_j \hat c_j \hat B_{d,j,\tau^\prime}$.

Using the regularized version of the theorem, we see that the coefficients $b_i=0$ for $i\leq d + 1$,
because they are a linear combination of $c_j=0$ for $j\leq d + 1$ by $\mu=d + 1$.
The same applies to the trailing coefficients, for which we would have $\mu = \# \tau^\prime$.
It follows that the projection from $\mathbb S_{d,\mathbf t^\prime}\to \mathbb S_{d,\mathbf t}$ preserves $f$.


\section*{Problem 4 (4.7)}
Let $1 = \sum B_{d,j,\tau}c_j$ with $c_j=1$ for all $j$.
Then $1 = \sum B_{d,i,\mathbf t} b_i$ with $b_i=\sum \alpha_{j,d}(i) c_j$,
however $\sum B_{d,i,\mathbf t}=1$, and the $B_{d,i,\mathbf t}$ are linearly independent,
so $b_i = 1$.
So $1=b_i = \sum\alpha_{j,d}(i)c_j=\sum\alpha_{j,d}(i)$.

\section*{Problem 5 (4.8)}
\lstinputlisting{src/test-oslo.m}
\lstinputlisting{src/oslo.m}
\lstinputlisting{src/computeR.m}
\lstinputlisting{output/oslo.txt}

\section*{Problem 6 (4.10)}
The blossoms of constant polynomials must be the polynomial itself,
so $\mathcal B[0](x_1,x_2,x_3) = 0$ and $\mathcal B[1](x_1,x_2,x_3) = 1$.

For monomials, the blossom is the sum of the symmetric homogeneous polynomials of the same degree,
so
\begin{align*}
    \mathcal B[x](x_1,x_2,x_3) 
    &= \frac{x_1 + x_2 + x_3} 3\\
    \mathcal B[x^2](x_1,x_2,x_3) 
    &= \frac{x_2x_3 + x_1x_3 + x_1x_2} 3\\
    \mathcal B[x](x_1,x_2,x_3) 
    &= x_1x_2x_3.
\end{align*}

The blossom is linear in it's polynomial,
so
\begin{equation*}\begin{aligned}
    \mathcal B[2x + x^2 - 4x^3](x_1,x_2,x_3) 
    &=2\mathcal B[x](x_1,x_2,x_3) + \mathcal B[x^2](x_1,x_2,x_3) - 4\mathcal B[x^3](x_1,x_2,x_3)\\
    &= 2(x_1 + x_2 + x_3) + (x_2x_3 + x_1x_3 + x_1x_2) - 4x_1x_2x_3.
\end{aligned}
\end{equation*}

By the linearity of the blossom, it must also be invariant under a linear change of base
\begin{equation*}\begin{aligned}
    \mathcal B[(x - a)^2](x_1,x_2,x_3)
    &= \mathcal B[x^2](x_1 - a,x_2 - a, x_3 - a)\\
    &= \frac{(x_2 - a)(x_3 - a) + (x_1 - a)(x_3 - a) + (x_1 - a)(x_2 - a)} 3
\end{aligned}
\end{equation*}

\end{document}
